---
title: "lettuce_growth_data"
author: "Sam H."
date: "2024-02-15"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
library(tidyverse)
library(ggplot2)
```

##Assigning a variable to the lettuce dataset
```{r}
lettuce_data <- read_csv("lettuce_dataset_1.5.csv", 
                         locale = readr::locale(encoding = "latin1"))
```
#This is an altered .csv file displaying data for only 1 of the 70
#specimens

#Alternatively, I could filter data down to each specimen with
#this code block and a pipe operator:
  
#lettuce_data %>%
# filter(plant_id = '1')

#but I decided it would be easier to alter the data range in the .csv itself

##Renaming col_names for plotting
```{r}
names(lettuce_data)[names(lettuce_data) == 'Temperature (C)'] <- 'temp'
names(lettuce_data)[names(lettuce_data) == 'Humidity (%)'] <- 'humidity'
names(lettuce_data)[names(lettuce_data) == 'TDS Value (ppm)'] <- 'tdsv'
names(lettuce_data)[names(lettuce_data) == 'pH Level'] <- 'ph_lvl'
names(lettuce_data)[names(lettuce_data) == 'Growth Days'] <- 'growth_days'
names(lettuce_data)[names(lettuce_data) == 'Plant_ID'] <- 'plant_id'
```

##Checking variable
glimpse(lettuce_data)
colnames(lettuce_data)

##Plotting data 

##Reviewing temperature trends
```{r}
ggplot(data = lettuce_data, aes(x = growth_days, y = temp)) +
  geom_point()+
  geom_smooth()+
  labs(title='Lettuce Growth: Temperature throughout Growth Days', 
           subtitle='Sample 1 of 70 individuals',
           caption= 'Data collected by Jjay Fabor',
           x= "Growth days",
           y= "Temperature (C)")
```

##Searching for patterns in humidity over growth period
```{r}
ggplot(data = lettuce_data, aes(x = growth_days, y = humidity)) +
  geom_point()+
  geom_smooth()+
  labs(title='Lettuce Growth: Humidity throughout Gorwth Days', 
           subtitle='Sample 1 of 70 individuals',
           caption= 'Data collected by Jjay Fabor',
           x= 'Growth days',
           y= 'Humidity') 
```

##Analyzing the relationship between pH level and TDS value
```{r}
lettuce_data %>%
  ggplot(aes(ph_lvl, tdsv))+
  geom_point(size = 3, alpha = 0.5)+
  geom_smooth(method = lm, se = F)+
  labs(title='Lettuce Growth: pH level vs. TDS value (ppm)', 
       subtitle='Sample 1 of 70 individuals',
       caption= 'Data collected by Jjay Fabor')
```



